Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-04 main.tex
:pencil: Add AO rate maximization

% amsthm
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

@@

		Back to the Pareto channel singular value problem \eqref{op:pareto}. The Euclidean gradient of \gls{bd}-\gls{ris} group $g$ is computed as
		\begin{equation}
			\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
			\label{eq:pareto_gradient}
		\end{equation}
		where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		Algorithm~\ref{ag:rcg} can thus be invoked for the Pareto singular value problem \eqref{op:pareto}, where expression \eqref{eq:pareto_gradient} should be used in line \ref{ln:gradient_euclidean}.
	\end{subsection}

@@

		There exist many singular value bounds on the product of two matrices in terms of their own singular values.
		One comprehensive answer is \cite{Fulton2000}
		\begin{equation}
			\prod_{k \in {K}} \sigma_k(\mathbf{H}) \le \prod_{i \in {I}} \sigma_i(\mathbf{H}_\mathrm{B}) \prod_{j \in {J}} \sigma_j(\mathbf{H}_\mathrm{F}),
			\label{iq:sv_bound_fc}
		\end{equation}
		for all admissible triples $(I, J, K) \in T_r^n$ with $r < n$, where
		\begin{equation}
			\begin{gathered}
				% T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \text{for all } p < r \text{ and all } (F, G, H) \text{ in } T_p^r,\\
				% \sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
				T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \forall p < r, (F, G, H) \in T_p^r,\\
				\sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
			\end{gathered}
		\end{equation}
		\begin{equation}
			U_r^n = \Bigl\{(I, J, K) \mid \sum_{i \in I} i + \sum_{j \in J} j = \sum_{k \in K} k + r(r+1)/2\Bigr\}.
		\end{equation}
		It implies the following special cases:
		\begin{itemize}
			\item upper bound on the largest singular value
			\begin{equation}
				\sigma_1(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item lower bound on the smallest singular value
			\begin{equation}
				\sigma_N(\mathbf{H}) \ge \sigma_N(\mathbf{H}_\mathrm{B}) \sigma_N(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item upper bound on the product of first $k$ singular values
			\begin{equation}
				\prod_{n=1}^k \sigma_n(\mathbf{H}) \le \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item lower bound on the product of last $k$ singular values
			\begin{equation}
				\prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}) \ge \prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item upper bound on the sum of first $k$ singular values to the power of $p > 0$
			\begin{equation}
				\sum_{n=1}^k \sigma_n^p(\mathbf{H}) \le \sum_{n=1}^k \sigma_n^p(\mathbf{H}_\mathrm{B}) \sigma_n^p(\mathbf{H}_\mathrm{F});
				\label{iq:sv_bound_fc_power}
			\end{equation}
			when $k = N$ and $p = 2$, it suggests the channel power is upper bounded by the sum of element-wise product of backward and forward subchannel power.
		\end{itemize}

		\begin{remark}
			Interestingly, \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when $\mathbf{\Theta} = \mathbf{V}_\mathrm{B} \mathbf{U}_\mathrm{F}^\mathsf{H}$.
			From \eqref{eq:scattering_fc} and \eqref{eq:channel_equivalent_fc}, we conclude the off-diagonal entries can enhance the capabilities of
			\begin{itemize}
				\item subspace alignment: $\mathbf{V}_\mathrm{B}$ and $\mathbf{U}_\mathrm{F}^\mathsf{H}$ in \eqref{eq:scattering_fc} fully align the subspaces of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ by rotation;
				\item subchannel rearrangement: $\mathbf{X} = \mathbf{I}$ in \eqref{eq:channel_equivalent_fc} pairs the components of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ by strength, which is optimal by rearrangement inequality.
			\end{itemize}
		\end{remark}

		Tight bounds are generally unavailable when the direct link exists, as the \gls{ris} needs to balance the backward-forward (multiplicative) and direct-indirect (additive) alignments.

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_singular_pareto.tex}
			}
			\caption{Singular value Pareto frontier. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:pc_singular_pareto}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_singular_bound.tex}
			}
			\caption{Singular value bounds for rank-1 indirect channel. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 32, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:pc_singular_bound}
		\end{figure}

		The Pareto frontier and evolving trend of channel singular values are shown in Fig.~\ref{sm:pc_singular_pareto} and \ref{sm:pc_singular_bound}.
		Clearly, \gls{bd}-\gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
	\begin{subsection}{Channel Power Maximization}
		Consider a \gls{bd}-\gls{ris} with $N^\mathrm{S}$ elements, which is divided into $G$ groups of equal $L$ elements.
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\left\lVert \mathbf{H}^\mathrm{D} + \sum_g\nolimits \mathbf{H}_g^\mathrm{B} \mathbf{\Theta}_g \mathbf{H}_g^\mathrm{F} \right\rVert _\mathrm{F}^2}{\label{op:pc_power}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g \in \mathcal{G} \triangleq \{1,\ldots,G\}.}{}{}
		\end{maxi!}
		For \emph{symmetric} BD-RIS, the problem has been solved in
		\begin{itemize}
			\item Matteo's paper \cite{Nerini2023}: SISO and equivalent\footnote{Single-stream MIMO with given precoder and combiner.};
			\item Ignacio's paper \cite{Santamaria2023}: SISO and directless MISO/SIMO.
		\end{itemize}

@@

	\begin{subsection}{Rate Maximization}
		The problem is formulated w.r.t. precoder (instead of transmit covariance matrix) for reference:
		\begin{maxi!}
			{\scriptstyle{\mathbf{W},\mathbf{\Theta}}}{R = \log \det \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\eta}\biggr)}{\label{op:pc_rate}}{\label{ob:pc_rate}}
			\addConstraint{\lVert \mathbf{W} \rVert _\mathrm{F}^2 \le P}{}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{\label{cs:pc_rate_unitary}}
		\end{maxi!}

		\eqref{op:pc_rate} is jointly non-convex and solved by \gls{ao}.
		For a given $\mathbf{\Theta}$, the optimal precoder is given by
		\begin{equation}
			\mathbf{W}^\star = \mathbf{V} {\mathbf{S}^\star}^{1/2},
		\end{equation}
		where $\mathbf{V}$ is right singular matrix of $\mathbf{H}$ and $\mathbf{S}^\star$ is a diagonal matrix of the water-filling power allocation.
		For a given $\mathbf{W}$,
		The complex derivative of \eqref{ob:pc_rate} w.r.t. \gls{ris} block $g$ is
		\begin{equation}
			\frac{\partial R}{\partial \mathbf{\Theta}_g^*} = \frac{1}{\eta} {\mathbf{H}_g^\mathrm{B}}^\mathsf{H} \mathbf{HW} \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\eta}\biggr)^{-1} \mathbf{W}^\mathsf{H} {\mathbf{H}_g^\mathrm{F}}^\mathsf{H}.
			\label{eq:pc_rate_gradient_ris}
		\end{equation}
		Algorithm~\ref{ag:pc_rate_ris} summarizes the adapted \gls{rcg} method for the \gls{ris} rate maximization subproblem.

		\setalgorithmcaptionfont{\small}
		\begin{algorithm}[!t]
			\small
			\caption{\gls{rcg} Method for \gls{ris} \gls{mimo}-\gls{pc} Rate Maximization}
			\label{ag:pc_rate_ris}
			\begin{algorithmic}[1]
				\Require $\mathbf{H}^\mathrm{D}$, $\mathbf{H}^\mathrm{F}$, $\mathbf{H}^\mathrm{B}$, $\mathbf{W}$, $L$, $\eta$
				\Ensure $\mathbf{\Theta}^\star$
				\State $r \gets 0$, $\mathbf{\Theta}^{(0)}$
				\Repeat
					\State $r \gets r+1$
					\For {$g \gets 1$ to $G$}
						\State $\mathbf{\Theta}_g^{(r)} \gets$ \eqref{eq:pc_rate_gradient_ris}, \eqref{eq:rcg_gradient_riemannian}--\eqref{eq:rcg_update}
					\EndFor
				\Until $\lvert R^{(r)} - R^{(r-1)} \rvert / R^{(r-1)} \le \epsilon$
			\end{algorithmic}
		\end{algorithm}

		\begin{figure}[!t]
			\centering
			\subfloat[\gls{ris} Elements, $N^\mathrm{T} = 8$\label{sm:pc_rate_sx}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_rate_sx.tex}
				}
			}
			\subfloat[Transmit Antenna, $N^\mathrm{S} = 256$\label{sm:pc_rate_tx}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/pc_rate_tx.tex}
				}
			}
			\caption{Average achievable rate versus group size $L$. $N^\mathrm{R} = 4$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
			\label{sm:pc_rate}
		\end{figure}

		Fig.~\ref{sm:pc_rate_sx} illustrates how \gls{ris} configuration influences the \gls{mimo} \gls{pc} achievable rate.
		To ensure a \qty{20}{bit/s/Hz} transmission, an \gls{snr} of \qty{13.5}{\dB} is required for a 8T4R system.
		This value decreases to \qty{12.5}{\dB} (resp. \qty{8}{\dB}) when 32- (resp. 256-) element diagonal \gls{ris} is present.
		If tetrads can be formed in \gls{bd}-\gls{ris}, the \gls{snr} can be reduced by another \qty{20}{\percent} (resp. \qty{44}{\percent}).
		Further increase in $L$ yields a marginal gain and incurs $\mathcal{O}(L^2)$ connections.
		We thus conclude dyadic or tetradic \gls{bd}-\gls{ris} usually strike a good balance between performance and complexity.
	\end{subsection}

>>>>

% amsthm
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}

@@

		% The Pareto singular value problem \eqref{op:pareto} can thus be solved by Algorithm~\ref{ag:rcg}.
		\begin{lemma}
			The Euclidean gradient of \eqref{ob:pareto} w.r.t. \gls{bd}-\gls{ris} group $g$ is
			\begin{equation}
				\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
				\label{eq:pareto_gradient}
			\end{equation}
			where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		\end{lemma}
		\begin{proof}
			Let $\mathbf{H} = \sum_n \mathbf{u}_n \sigma_n \mathbf{v}_n^\mathsf{H}$ be the compact \gls{svd} of the equivalent channel.
			Since the singular vectors are orthonormal, the $n$-th singular value can be expressed as
			\begin{equation}
				\sigma_n = \mathbf{u}_n^\mathsf{H} \mathbf{H} \mathbf{v}_n = \mathbf{u}_n^\mathsf{T} \mathbf{H}^* \mathbf{v}_n^*,
			\end{equation}
			whose differential w.r.t. $\mathbf{\Theta}_g^*$ is
			\begin{align*}
				\partial \sigma_n
				& = \partial \mathbf{u}_n^\mathsf{T} \underbrace{\mathbf{H}^* \mathbf{v}_n^*}_{\sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T} \mathbf{v}_n} + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \underbrace{\mathbf{u}_n^\mathsf{T} \mathbf{H}^*}_{\mathbf{u}_n^\mathsf{T} \sum_m \mathbf{u}_m^* \sigma_m \mathbf{v}_m^\mathsf{T}} \partial \mathbf{v}_n^*\\
				& = \underbrace{\partial \mathbf{u}_n^\mathsf{T} \mathbf{u}_n^*}_{\partial 1 = 0} \cdot \sigma_n + \mathbf{u}_n^\mathsf{T} \cdot \partial \mathbf{H}^* \cdot \mathbf{v}_n^* + \sigma_n \cdot \underbrace{\mathbf{v}_n^\mathsf{T} \partial \mathbf{v}_n^*}_{\partial 1 = 0}\\
				& = \mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \cdot \partial \mathbf{\Theta}_g^* \cdot \mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\\
				& = \mathrm{tr}(\mathbf{H}_{\mathrm{F},g}^* \mathbf{v}_n^*\mathbf{u}_n^\mathsf{T} \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \cdot \partial \mathbf{\Theta}_g^*).
			\end{align*}
			According to \cite{Hjorungnes2007}, the corresponding complex derivative is
			\begin{equation}
				\frac{\partial \sigma_n}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{u}_n \mathbf{v}_n^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
				\label{eq:derivative_sv}
			\end{equation}
			Linear combination of \eqref{eq:derivative_sv} yields \eqref{eq:pareto_gradient}.
		\end{proof}

		%  The Euclidean gradient of \gls{bd}-\gls{ris} group $g$ is computed as
		% where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		Algorithm~\ref{ag:rcg} can thus be invoked for the Pareto singular value problem \eqref{op:pareto}, where line \ref{ln:gradient_euclidean} uses explicit expression \eqref{eq:pareto_gradient}.
		% \eqref{eq:pareto_gradient} is used in line \ref{ln:gradient_euclidean}.
	\end{subsection}

@@

		There exist many singular value bounds on the product of two matrices in terms of their own singular values.
		One comprehensive answer is \cite{Fulton2000}
		\begin{equation}
			\prod_{k \in {K}} \sigma_k(\mathbf{H}) \le \prod_{i \in {I}} \sigma_i(\mathbf{H}_\mathrm{B}) \prod_{j \in {J}} \sigma_j(\mathbf{H}_\mathrm{F}),
			\label{iq:sv_bound_fc}
		\end{equation}
		for all admissible triples $(I, J, K) \in T_r^n$ with $r < n$, where
		\begin{equation}
			\begin{gathered}
				% T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \text{for all } p < r \text{ and all } (F, G, H) \text{ in } T_p^r,\\
				% \sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
				T_r^n \triangleq \Bigl\{(I, J, K) \in U_r^n \mid \forall p < r, (F, G, H) \in T_p^r,\\
				\sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
			\end{gathered}
		\end{equation}
		\begin{equation}
			U_r^n \triangleq \Bigl\{(I, J, K) \mid \sum_{i \in I} i + \sum_{j \in J} j = \sum_{k \in K} k + r(r+1)/2\Bigr\}.
		\end{equation}
		It implies the following special cases:
		\begin{itemize}
			\item upper bound on the largest singular value
			\begin{equation}
				\sigma_1(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item lower bound on the smallest singular value
			\begin{equation}
				\sigma_N(\mathbf{H}) \ge \sigma_N(\mathbf{H}_\mathrm{B}) \sigma_N(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item upper bound on the product of first $k$ singular values
			\begin{equation}
				\prod_{n=1}^k \sigma_n(\mathbf{H}) \le \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item lower bound on the product of last $k$ singular values
			\begin{equation}
				\prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}) \ge \prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=N}^{N-k+1} \sigma_n(\mathbf{H}_\mathrm{F});
			\end{equation}
			\item upper bound on the sum of first $k$ singular values to the power of $p > 0$
			\begin{equation}
				\sum_{n=1}^k \sigma_n^p(\mathbf{H}) \le \sum_{n=1}^k \sigma_n^p(\mathbf{H}_\mathrm{B}) \sigma_n^p(\mathbf{H}_\mathrm{F});
				\label{iq:sv_bound_fc_power}
			\end{equation}
			when $k = N$ and $p = 2$, it suggests the channel power is upper bounded by the sum of (sorted) element-wise power product of backward and forward subchannel.
		\end{itemize}

		\begin{remark}
			Interestingly, \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when $\mathbf{\Theta} = \mathbf{V}_\mathrm{B} \mathbf{U}_\mathrm{F}^\mathsf{H}$.
			From \eqref{eq:scattering_fc} and \eqref{eq:channel_equivalent_fc}, we conclude the off-diagonal entries can enhance the capabilities of
			\begin{itemize}
				\item subspace alignment: $\mathbf{V}_\mathrm{B}$ and $\mathbf{U}_\mathrm{F}^\mathsf{H}$ in \eqref{eq:scattering_fc} fully align the subspaces of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ by rotation;
				\item subchannel rearrangement: $\mathbf{X} = \mathbf{I}$ in \eqref{eq:channel_equivalent_fc} pairs the subchannels of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ from strongest to weakest, which is optimal by rearrangement inequality.
			\end{itemize}
		\end{remark}

		Tight bounds are generally unavailable when \gls{mimo} direct link is present.
		This is because the \gls{ris} needs to balance the backward-forward (multiplicative) and direct-indirect (additive) subspace alignments.

		% \begin{figure}[!t]
		% 	\centering
		% 	\resizebox{0.65\columnwidth}{!}{
		% 		\input{assets/simulation/pc_singular_pareto.tex}
		% 	}
		% 	\caption{Singular value Pareto frontier. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
		% 	\label{sm:pc_singular_pareto}
		% \end{figure}

		% \begin{figure}[!t]
		% 	\centering
		% 	\resizebox{0.65\columnwidth}{!}{
		% 		\input{assets/simulation/pc_singular_bound.tex}
		% 	}
		% 	\caption{Singular value bounds for rank-1 indirect channel. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 32, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
		% 	\label{sm:pc_singular_bound}
		% \end{figure}

		% The Pareto frontier and evolving trend of channel singular values are shown in Fig.~\ref{sm:pc_singular_pareto} and \ref{sm:pc_singular_bound}.
		% Clearly, \gls{bd}-\gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}

\begin{section}{Rate Maximization}
	The \gls{mimo} rate maximization problem is formulated w.r.t. joint active and passive beamforming
	\begin{maxi!}
		{\scriptstyle{\mathbf{W},\mathbf{\Theta}}}{R = \log \det \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\eta}\biggr)}{\label{op:rate}}{\label{ob:rate}}
		\addConstraint{\lVert \mathbf{W} \rVert _\mathrm{F}^2 \le P}{}{}
		\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
	\end{maxi!}
	where $\mathbf{W}$ is the transmit precoder, $R$ is the achievable rate, $\eta$ is the noise power, and $P$ is the transmit power budget.
	Two methods are proposed below to solve problem \eqref{op:rate}.

	\begin{subsection}{Alternating Optimization}
		We first consider an \gls{ao} approach that updates $\mathbf{\Theta}$ and $\mathbf{W}$ iteratively.
		For a given $\mathbf{W}$, the passive beamforming subproblem is
		\begin{maxi!}
			% {\scriptstyle{\mathbf{\Theta}}}{\log \det \biggl(\mathbf{I} + \frac{(\mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F}) \mathbf{Q} (\mathbf{H}_\mathrm{D}^\mathsf{H} + \mathbf{H}_\mathrm{F}^\mathsf{H} \mathbf{\Theta}^\mathsf{H} \mathbf{H}_\mathrm{B}^\mathsf{H})}{\eta}\biggr)}{\label{op:rate_passive}}{\label{ob:rate_passive}}
			{\scriptstyle{\mathbf{\Theta}}}{\log \det \biggl(\mathbf{I} + \frac{\mathbf{H} \mathbf{Q} \mathbf{H}^\mathsf{H}}{\eta}\biggr)}{\label{op:rate_passive}}{\label{ob:rate_passive}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		\end{maxi!}
		where $\mathbf{Q} \triangleq \mathbf{W} \mathbf{W}^\mathsf{H}$ is the transmit covariance matrix.
		\begin{lemma}
			The Euclidean gradient of \eqref{ob:rate_passive} w.r.t. \gls{bd}-\gls{ris} block $g$ is
			\begin{equation}
				\frac{\partial R}{\partial \mathbf{\Theta}_g^*} = \frac{1}{\eta} \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \biggl(\mathbf{I} + \frac{\mathbf{H}\mathbf{Q}\mathbf{H}^\mathsf{H}}{\eta}\biggr)^{-1} \mathbf{H} \mathbf{Q} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
				\label{eq:rate_gradient}
			\end{equation}
		\end{lemma}

		\begin{proof}
			The differential of $R$ w.r.t. $\mathbf{\Theta}_g^*$ is \cite{Hjorungnes2007}
			\begin{align*}
				\partial R
				& = \frac{1}{\eta} \mathrm{tr} \biggl\{ \partial \mathbf{H}^* \cdot \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T} \Bigl(\mathbf{I} + \frac{\mathbf{H}^* \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T}}{\eta}\Bigr)^{-1} \biggr\}\\
				& = \frac{1}{\eta} \mathrm{tr} \biggl\{ \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^* \cdot \mathbf{H}_{\mathrm{F},g}^* \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T} \Bigl(\mathbf{I} + \frac{\mathbf{H}^* \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T}}{\eta}\Bigr)^{-1} \biggr\}\\
				& = \frac{1}{\eta} \mathrm{tr} \biggl\{ \mathbf{H}_{\mathrm{F},g}^* \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T} \Bigl(\mathbf{I} + \frac{\mathbf{H}^* \mathbf{Q}^\mathsf{T} \mathbf{H}^\mathsf{T}}{\eta}\Bigr)^{-1} \mathbf{H}_{\mathrm{B},g}^* \cdot \partial \mathbf{\Theta}_g^* \biggr\},
			\end{align*}
			and the corresponding complex derivative is \eqref{eq:rate_gradient}.
		\end{proof}
		Problem \eqref{op:rate_passive} is then solved by Algorithm \ref{ag:rcg}, where line \ref{ln:gradient_euclidean} uses explicit expression \eqref{eq:rate_gradient}.
		For a given $\mathbf{\Theta}$, the optimal transmit precoder is given by eigenmode transmission \cite{Clerckx2013}
		\begin{equation}
			\mathbf{W}^\star = \mathbf{V} {\mathbf{S}^\star}^{1/2},
		\end{equation}
		where $\mathbf{V}$ is the right singular matrix of $\mathbf{H}$ and $\mathbf{S}^\star$ is the optimal power allocation matrix obtained by the water-filling algorithm.
	\end{subsection}

	\begin{subsection}{Low-Complexity Solution}

	\end{subsection}

	% \begin{figure}[!t]
	% 	\centering
	% 	\subfloat[\gls{ris} Elements, $N^\mathrm{T} = 8$\label{sm:pc_rate_sx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/simulation/pc_rate_sx.tex}
	% 		}
	% 	}
	% 	\subfloat[Transmit Antenna, $N^\mathrm{S} = 256$\label{sm:pc_rate_tx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/simulation/pc_rate_tx.tex}
	% 		}
	% 	}
	% 	\caption{Average achievable rate versus group size $L$. $N^\mathrm{R} = 4$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{sm:pc_rate}
	% \end{figure}

	% Fig.~\ref{sm:pc_rate_sx} illustrates how \gls{ris} configuration influences the \gls{mimo} \gls{pc} achievable rate.
	% To ensure a \qty{20}{bit/s/Hz} transmission, an \gls{snr} of \qty{13.5}{\dB} is required for a 8T4R system.
	% This value decreases to \qty{12.5}{\dB} (resp. \qty{8}{\dB}) when 32- (resp. 256-) element diagonal \gls{ris} is present.
	% If tetrads can be formed in \gls{bd}-\gls{ris}, the \gls{snr} can be reduced by another \qty{20}{\percent} (resp. \qty{44}{\percent}).
	% Further increase in $L$ yields a marginal gain and incurs $\mathcal{O}(L^2)$ connections.
	% We thus conclude dyadic or tetradic \gls{bd}-\gls{ris} usually strike a good balance between performance and complexity.

	\begin{subsection}{Channel Power Maximization}
		Consider a \gls{bd}-\gls{ris} with $N^\mathrm{S}$ elements, which is divided into $G$ groups of equal $L$ elements.
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\left\lVert \mathbf{H}^\mathrm{D} + \sum_g\nolimits \mathbf{H}_g^\mathrm{B} \mathbf{\Theta}_g \mathbf{H}_g^\mathrm{F} \right\rVert _\mathrm{F}^2}{\label{op:pc_power}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g \in \mathcal{G} \triangleq \{1,\ldots,G\}.}{}{}
		\end{maxi!}
		For \emph{symmetric} BD-RIS, the problem has been solved in
		\begin{itemize}
			\item Matteo's paper \cite{Nerini2023}: SISO and equivalent\footnote{Single-stream MIMO with given precoder and combiner.};
			\item Ignacio's paper \cite{Santamaria2023}: SISO and directless MISO/SIMO.
		\end{itemize}

@@
